---
title: "STAT 545A - Exploring Datasets"
author: "Aidan Hughes"
output:
  html_document:
    theme: cerulean
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the file, we're going try playing around with a couple of datasets.

##Let's get it started

We're going to start with the Gapminder data frame to try out some R functions.

```{r}
library(gapminder)

head(gapminder)
```

So it looks like we have a bunch of data on some countries. Let's see if we have anything on our favorite countries, Canada and the US.

```{r}
any(gapminder=="Canada")
any(gapminder=="United States")
```

Great, now lets compare the two in a nice plot.

```{r, echo=FALSE}
CAN_data <- gapminder[gapminder$country == "Canada",]
US_data <- gapminder[gapminder$country == "United States",]

# Define identical Y limits for the datasets
ylims <- c(50,90)

plot(CAN_data$year, CAN_data$lifeExp, ylim=ylims, col="blue", pch=19, ylab="Life Expentancy (years)", xlab="Date")
par(new=TRUE)
plot(US_data$year, US_data$lifeExp, ylim=ylims, ylab="", xlab="", col="red", pch=19)
legend(x=1955,y=90, legend=c("Canada", "United States"),col=c("red", "blue"), fill=c("red", "blue"))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
