---
title: "STAT 545A - Exploring Datasets"
author: "Aidan Hughes"
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the file, we're going try playing around with a couple of datasets. We'll start with Gapminder like we did in class, and then move onto some of the built-in datasets in R that are available to us.

## Let's Get Started - Gapminder

We're going to start with the Gapminder data frame to try out some R functions.

```{r echo=FALSE}
library(gapminder)

```

```{r}
head(gapminder)
```

So it looks like we have a bunch of data on some countries. Let's see if we have anything on our favorite countries, Canada and the US, using the `any()` function.

```{r}
any(gapminder=="Canada")
any(gapminder=="United States")
```

Great, now lets compare the two in a nice plot.

```{r, echo=FALSE}
CAN_data <- gapminder[gapminder$country == "Canada",]
US_data <- gapminder[gapminder$country == "United States",]

# Define identical Y limits for the datasets
ylims <- c(50,90)

plot(CAN_data$year, CAN_data$lifeExp, ylim=ylims, col="blue", pch=19,main="Life Expectancy of Canada vs. The United States", ylab="Life Expentancy (years)", xlab="Date")
par(new=TRUE)
plot(US_data$year, US_data$lifeExp, ylim=ylims, ylab="", xlab="", col="red", pch=19)
legend(x=1955,y=90, legend=c("Canada", "United States"),col=c("red", "blue"), fill=c("red", "blue"), bg="lightgrey")
```

Well, I was hoping Canada would do better than the States in this category but I guess not. Let's move on, shall we?

## Built-in Datasets

There's plenty of built-in datasets meant just for this purpose - getting used to R. 
Because I found out how to do it on the internet, let's plot the surface graph of a volcano!

```{r}
#--- This code block renders a 3D surface plot of a volcano, but unfortunately only works with output as HTML so I have omitted it. #--- It also adds about 4mb to the file. It does work though, take my word.

#library(plotly)
#p <- plot_ly(z = ~volcano) %>% add_surface()
#p
```

### Well, I hope this is enough for a reasonable grade. Thanks for your time!